---
output:
  word_document: default
  html_document: default
---
#This is My Markdown File
Siddhesh Dalavi

##Reproducible search, week2 assignment
```{r setoptions,echo=TRUE}
library(knitr)
opts_chunk$set(echo=TRUE,results="asis")
```

Load the data.
```{r simulation}
Activity_data<-read.csv("activity.csv")
summary(Activity_data)
head(Activity_data)
```

Remove missing values
```{r}
act.complete<-na.omit(Activity_data)
head(act.complete)
```

Total Number of steps taken
```{r}
library(dplyr)
act.day<-group_by(act.complete,date)
act.day<-summarize(act.day,steps=sum(steps))
head(act.day)
```

Plot the histogram
```{r}
library(ggplot2)
qplot(steps,data=act.day,xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

calculating the mean and median of steps taken per day
```{r}
mean(act.day$steps)
median(act.day$steps)
```

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
act.int<-group_by(act.complete,interval)
act.int<-summarize(act.int,steps=mean(steps))
```

plot the graph
```{r}
ggplot(act.int,aes(interval,steps))+geom_line()
```

We find the row in the interval data frame for which steps is equal to the maximum number of steps, then we look at the interval of that row
```{r}
act.int[act.int$steps==max(act.int$steps),]
```

total number of NA values
```{r}
nrow(Activity_data)-nrow(act.complete)
```

filling up all missing values
```{r}
names(act.int)[2] <- "mean.steps"
act.impute <- merge(Activity_data, act.int)
```

create new dataset with filled values
```{r}
act.impute$steps[is.na(act.impute$steps)] <- act.impute$mean.steps[is.na(act.impute$steps)]
```

create a dataset with imputed values
```{r}
act.day.imp <- group_by(act.impute, date)
act.day.imp <- summarize(act.day.imp, steps=sum(steps))
```

plot the graph 
```{r}
qplot(steps, data=act.day.imp)
mean(act.day.imp$steps)
median(act.day.imp$steps)
```

Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
act.impute$dayofweek <- weekdays(as.Date(act.impute$date))
act.impute$weekend <-as.factor(act.impute$dayofweek=="Saturday"|act.impute$dayofweek=="Sunday")
levels(act.impute$weekend) <- c("Weekday", "Weekend")
```

Create seperate dataframes for weekday and weekend, also take mean for steps
```{r}
act.weekday <- act.impute[act.impute$weekend=="Weekday",]
act.weekend <- act.impute[act.impute$weekend=="Weekend",]
act.int.weekday <- group_by(act.weekday, interval)
act.int.weekday <- summarize(act.int.weekday, steps=mean(steps))
act.int.weekday$weekend <- "Weekday"
act.int.weekend <- group_by(act.weekend, interval)
act.int.weekend <- summarize(act.int.weekend, steps=mean(steps))
act.int.weekend$weekend <- "Weekend"
```

now append the 2 dataframes and plot the time series
```{r}
act.int <- rbind(act.int.weekday, act.int.weekend)
act.int$weekend <- as.factor(act.int$weekend)
ggplot(act.int, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .)
```









